---
title: "bobcat_conflict"
author: "Lara Volski"
date: "11/11/2021"
output: html_document
---
# Part 1) Setting up

### 1.1 Load Packages
```{r library.packages, include = FALSE}
library(nlme)
library(ggplot2)
library(overlap)
library(maptools)
library(lubridate)
library(plyr)
library(camtrapR)
library(dplyr)
library(lme4)
library(MASS)
library(circular) #Watson2Test
library(ggpattern)
```

### 1.2 Read in record table CSV
```{r record.table}
record.table.all <- read.csv("record_table_0min_deltaT_2020-06-06.csv")
```

### 1.3 create new column for cameras by taking the folder name and reducing to first three characters
```{r new.columns}
record.table.all$Camera <- strtrim(record.table.all$Camera_folder, 5)

record.table.all$Phase <- substr(record.table.all$Camera_folder, 7, 12)

record.table.all$Foxlight <- substr(record.table.all$Camera_folder, 14, 17)
```

### 1.4 Set the time interval (photos that occured within 15 min will count as one detection)
```{r fifteen.min}
record.table.15min <- record.table.all[record.table.all$delta.time.secs >= 900 |  # because 900 seconds = 15 minutes
                                             record.table.all$delta.time.secs == 0, ]
```

### 1.5 Import additional csv

*metadata.raster is normally distributed. The closer to 0, the less the value deviates from the norm. 
```{r import.csv}
# import camera phase operation dates
cam_operations <- read.csv("phase_operation_dates_2020.csv", header=T)

# reading in metadata

# metadata <- read.csv("MicrohabitatData.csv", header=T)
metadata.raster <- read.csv("Raster_metadata_Lara.csv", header=T)
```

### 1.6 Plotting Species Richness across cameras by lat and long
```{r species.richness.plot}
detectionMaps(CTtable = cam_operations, 
              recordTable = record.table.15min, 
              stationCol = "Camera", 
              Xcol = "Latitude", 
              Ycol = "Longitude",
              richnessPlot = TRUE,
              printLabels = TRUE)

# remove species that we don't care about (here, removing birds, opposum, nothing, mice, pigs, and squirrels) 
for (species in c("2", "Bird", "DIVI", "Ghosts", "Mice", "SCOC", "SCGR", "SCNI")) {
  record.table.15min <- record.table.15min[record.table.15min$Species != species, ]
}
```

### RAI
```{r RAI}
# calculate RAI for each period

# change camera, phase, and species to factor
record.table.15min$Camera <- as.factor(record.table.15min$Camera)
record.table.15min$Phase <- as.factor(record.table.15min$Phase)
record.table.15min$Species <- droplevels(as.factor(record.table.15min$Species))


# calculate number of observations of each species at each camera in each phase
rai <- record.table.15min %>%
  dplyr::group_by(Species, Camera, Phase, .drop = FALSE) %>%
  dplyr::summarise(Detections = n()) 

rai2 <- record.table.15min %>%
  dplyr::group_by(Species, .drop = FALSE) %>%
  dplyr::summarise(Detections = n())
# merge with record table
rai <- merge(rai, cam_operations)


# calculate RAI
rai$RAI <- rai$Detections / rai$Operation

# remove records where camera was operating for <10 days
for (i in 1:nrow(rai)) {
  if(rai$Operation[i] < 10) {
    rai$Detections[i] <- NA
    rai$RAI[i] <- NA
  } 
}

# join raster metadata
rai <- left_join(rai, metadata.raster)

#### now just calculate RAI at each camera, ACROSS ALL PHASES
cam_operations.acrossphases <- read.csv("camera_operation_dates_2020.csv")
# calculate number of observations of each species at each camera
rai.acrossphases <- record.table.15min %>%
  dplyr::group_by(Species, Camera, .drop = FALSE) %>%
  dplyr::summarise(Detections = n()) 
# merge with record table
rai.acrossphases <- merge(rai.acrossphases, cam_operations.acrossphases)
# calculate RAI
rai.acrossphases$RAI <- rai.acrossphases$Detections / rai.acrossphases$Operation
# join with metadata
rai.acrossphases <- left_join(rai.acrossphases, metadata.raster)
```

### Changing phases so that they correspond to FoxY1 and FoxN1.
```{r}
#Phase 1 and 2 --> Phase 1
# Phase 3 and 4 --> Phase 2
# Phase 5 and 6 --> Phase 3
# Phase 7 and 8 --> Phase 4
# Phase 9 --> Phase 5
phase_key <- tibble(Phase = c("Phase1", "Phase2", "Phase3", "Phase4", "Phase5", "Phase6", "Phase7", "Phase8", "Phase9"), Phase_New = c(1, 1, 2, 2, 3, 3, 4, 4, 5))

# Then use left_join to join this with the RAI data frame used for modeling, and use this column (Phase_New) as your covariate in modeling. 
rai <- left_join(rai, phase_key)
```
# Question 1: How do bobcat detections vary by foxlight presence?

## Test if RAI changes between foxlight yes and foxlight no, while controlling for camera

Bobcats
```{r q1.bobcat}

# The mult_format multiplies the coyote RAI by 35, so instead of looking at daily average RAI, we are looking at it per 5 week phase (because 7 days x 5 weeks = 35 days). If you want to take it to a 10 week phase, multiply it by 50.


mult_format <- function() {
     function(x) format(35*x,digits = 2)}


ggplot(subset(rai, Species == "LYRU"), aes(x = Foxlight, y = RAI*35, fill = Foxlight)) +
  geom_boxplot(alpha=2) +
  theme_bw() +
  theme(legend.position = "none") +
  scale_x_discrete(labels = c("No", "Yes")) +
 # geom_jitter() +
  labs(x ="Foxlight", y = "Relative Bobcat Activity (detections/phase)") +
  ## scale_y_continuous(labels = mult_format()) +
  scale_fill_manual(values=c("cadetblue4", "orange1")) + 
  ggsave("figures/bobcat_foxlight.png", width = 4, height = 4)

# Line Graph for Coyote RAI by Foxlight Presence and Camera Site
ggplot(subset(rai, Species == "LYRU"), aes(x = Phase, y = RAI, col = Foxlight, group = 1)) +
  geom_point() + geom_line(color="black") +
  facet_wrap(~ Camera) +
  scale_x_discrete(label=function(x) abbreviate(x, minlength=1)) +
  labs(title = "LYRU RAI by Foxlight Presence and Camera Site") +
  ggsave("figures/coyote_RAI_by_camera_site.png", width = 6, height = 4)
            
```

# Question 2: How do bobcat detections vary by sheep presence?

## Test if RAI changes between foxlight yes and foxlight no, while controlling for camera
```{r q2.bobcat}

# The mult_format multiplies the coyote RAI by 35, so instead of looking at daily average RAI, we are looking at it per 5 week phase (because 7 days x 5 weeks = 35 days). If you want to take it to a 10 week phase, multiply it by 50.

rai_bobcat <- read.csv("rai_bobcat.csv", header=T)

mult_format <- function() {
     function(x) format(35*x,digits = 2)}

## just sheep presence

ggplot(subset(rai_bobcat, Species == "LYRU"), aes(x = Sheep_Presence, y = RAI*35, fill = Sheep_Presence)) +
  geom_boxplot(alpha=2) +
  theme_bw() +
  theme(legend.position = "none") +
  scale_x_discrete(labels = c("No", "Yes")) +
 # geom_jitter() +
  labs(x ="Sheep Present", y = "Relative Bobcat Activity (detections/phase)") +
  ## scale_y_continuous(labels = mult_format()) +
  scale_fill_manual(values=c("cadetblue4", "orange1")) + 
  ggsave("figures/bobcat_sheep_presence.png", width = 4, height = 4)

## sheep presence with a Foxlight fill

ggplot(subset(rai_bobcat, Species == "LYRU"), aes(x = Sheep_Presence, y = RAI*35, fill = Foxlight)) +
  geom_boxplot(alpha=2) +
  theme_bw() +
  theme(legend.position = "none") +
  scale_x_discrete(labels = c("No", "Yes")) +
 # geom_jitter() +
  labs(x ="Sheep Presence", y = "Relative Bobcat Activity (detections/phase)") +
  ## scale_y_continuous(labels = mult_format()) +
  scale_fill_manual(values=c("cadetblue4", "orange1")) + 
  ggsave("figures/bobcat_sheep_presence_with_Foxlight.png", width = 4, height = 4)
            
```

# Question 3: How do bobcat detections vary by dog presence?

```{r}

##looking at dog RAI... likely double-counting the same dog...
rai_bobcat2 <- read.csv("rai_bobcat_lv.csv", header=T)

ggplot(data = rai_bobcat2, aes(x = Dog_RAI, y = RAI)) +
  geom_point(aes(col=Foxlight)) +
  geom_smooth(method="loess", se=F) +
  labs(x ="Dog RAI", y = "Bobcat RAI")

## Individual Dog Count
# geom_smooth better for two continuous variables, whereas here we have a discrete & continuous
ggplot(data = rai_bobcat2, aes(x = Ind_Dogs, y = RAI)) +
  geom_point(aes(col=Foxlight)) +
  geom_smooth(method="loess", se=F) +
  labs(x ="Dogs", y = "Bobcat RAI")

ggplot(data = subset(rai_bobcat2, Species == "LYRU"), aes(x = Ind_Dogs, y = Detections)) +
  geom_col(fill = "cadetblue4") +
  labs(x = "Dogs", y = "Bobcat Detections") +
  ggsave("figures/bobcat_col_individualdog.png", width = 4, height = 4)

ggplot(data = rai_bobcat2, aes(x = Ind_Dogs, y = RAI * 35, group = Ind_Dogs)) +
  geom_boxplot(alpha=2, fill = "cadetblue4") +
  labs(x ="Dogs", y = "Bobcat RAI") +
  theme(legend.position = "none") +
  ggsave("figures/bobcat_boxplot_individualdogs.png", width = 4, height = 4)


## Testing for p-values
t.test(sample.data$deer_before, sample.data$deer_during)

## try 0 vs. 1, 1 vs. 2, etc. to look for threshold.

```

# Question 4: Habituation
```{r q4.habituation}
##read in the CSV ##

habituation.by.rai.foxY <- read.csv("lyru.habituation.foxlights.csv", header=T)

ggplot(habituation.by.rai.foxY, aes(x = Phase, y = LYRU.RAI)) +
  geom_boxplot(fill = "cadetblue4") +
  labs(title = "LYRU RAI by Active Foxlight Phase",
       x = "First, Second, Third, Fourth, or Fifth Time a Foxlight was Active at a Site", y = "Bobcat RAI") +
  ggsave("figures/bobcat_foxlight_habituation.png", width = 6, height = 4)

medians.by.week <- read.csv("lyru.medians.csv", header=T)

ggplot(data = medians.by.week, aes(x = Week, y = Median, fill = Foxlight)) +
  geom_bar(stat = "identity", position = position_dodge())  +
  scale_x_continuous(breaks = seq(1, 9, by = 1)) +
  theme(plot.title = element_text(hjust = 0.5), 
        axis.title.x = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        legend.title = element_text(face="bold", size = 10)) +
labs(x ="Week", y = "Median Bobcat Detections") +
  scale_fill_manual(values=c("cadetblue4", "orange1")) + 
  theme_bw() +
  ggsave("figures/bobcat_weekly_detections.png", width = 4, height = 4)
```

# Question 5: Diel Patterns and Dogs
```{r q5.dailyactivity.bobcat}

## Only included LGDs for this. There were some detections of dogs that weren't LGDs. 

record.table.lyru <- read.csv("record.table.15min1.csv")

# Setting up
# specify date format
record.table.lyru$Date <- as.POSIXct(record.table.lyru$Date)

# scale clock time to solar time
coords <- matrix(c(-123.079, 39.0013), nrow=1) # note it is c(longitude, latitude)
Coords <- sp::SpatialPoints(coords,
                            proj4string=sp::CRS("+proj=longlat +datum=WGS84"))

# store time as hms
record.table.lyru$Time.Corrected <- hms(record.table.lyru$Time)

# convert time from HH:MM:SS to decimal hours (HH.HHHH or whatever)
record.table.lyru$Time.Decimal <- record.table.lyru$Time.Corrected$hour + record.table.lyru$Time.Corrected$minute/60 + record.table.lyru$Time.Corrected$second/3600

# scale time so that it is between 0 and 1 by dividing by 24
record.table.lyru$Time.Scaled <- record.table.lyru$Time.Decimal / 24

# convert scaled time (0 to 1) to radians (0 to 2pi)
record.table.lyru$Time.Radians <- record.table.lyru$Time.Scaled * 2 * pi

# calculate suntime using function from overlap package, and coordinates and dates as formatted above
record.table.lyru$Time.Sun <- sunTime(record.table.lyru$Time.Radians, record.table.lyru$Date, Coords)

# plot coyote activity with and without foxlights
records.bobcat.dogyes <- subset(record.table.lyru, Species == "LYRU" & Dog_Presence == "Y")
records.bobcat.dogno <- subset(record.table.lyru, Species == "LYRU" & Dog_Presence == "N")

# define overlap function
multilines2 <-function (A, B, xscale = 24, linetype = c(1, 2), linecol = c("orange1", "cadetblue4"), n.grid = 128, kmax = 3, adjust = 1,
                        ...) 
{
  bwA <- getBandWidth(A, kmax = kmax)/adjust
  bwB <- getBandWidth(B, kmax = kmax)/adjust
  if (is.na(bwA) || is.na(bwB)) 
    stop("Bandwidth estimation failed.")
  xsc <- if (is.na(xscale))
    1
  else xscale/(2 * pi)
  xxRad <- seq(0, 2 * pi, length = n.grid)
  xx <- xxRad * xsc
  densA <- densityFit(A, xxRad, bwA)/xsc
  densB <- densityFit(B, xxRad, bwB)/xsc
  densOL <- pmin(densA, densB)
  ylim <- c(0, max(densA, densB))
  plot(0, 0, type = "n", ylim = ylim, xlim = range(xx), xlab = "Time", 
       ylab = "Relative Bobcat Activity", xaxt = "n", ...)
  if (is.na(xscale)) {
    axis(1, at = c(0, pi/2, pi, 3 * pi/2, 2 * pi), labels = c("0", 
                                                              expression(pi/2), expression(pi), expression(3 * 
                                                                                                             pi/2), expression(2 * pi)))
  }
  else if (xscale == 24) {
    axis(1, at = c(0, 6, 12, 18, 24), labels = c("Midnight", 
                                                 "Sunrise", "Noon", "Sunset", "Midnight"))
  }
  else {
    axis(1)
  }
  lines(xx, densA, lty = linetype[1], col = linecol[1], lwd = 2)
  lines(xx, densB, lty = linetype[2], col = linecol[2], lwd = 2)
  return(invisible(list(x = xx, densityA = densA, densityB = densB)))
}

watson.two.test(records.bobcat.dogyes$Time.Sun, records.bobcat.dogno$Time.Sun)


# plot bobcat activity with and without dogs
records.bobcat.dogyes <- subset(record.table.lyru, Species == "LYRU" & Dog_Presence == "Y")
records.bobcat.dogno <- subset(record.table.lyru, Species == "LYRU" & Dog_Presence == "N")

# plot for bobcats
multilines2(records.bobcat.dogyes$Time.Sun, records.bobcat.dogno$Time.Sun, main = "Bobcats w/ and w/o Dog Presence")
legend('top', c("Yes Dogs", "No Dogs"), col = c("orange1", "cadetblue4"), lty=c(1,1), bty='n')

#test significance
# this will return the degree of overlap - area under the shared curve (where 0 is completely different and 1 is completely the same)
(Dhats <- overlapEst(records.bobcat.dogyes$Time.Sun, records.bobcat.dogno$Time.Sun))
# use Dhat4 (they are different formulas - this one is good for sample sizes >50)

# then resample each one 10,000 times (bootstrapping)
bs_time.dogyes <- resample(records.bobcat.dogyes$Time.Sun, 10000)
bs_time.dogno <- resample(records.bobcat.dogno$Time.Sun, 10000)

# now calculate the resulting bootstrapped overlap value
# it should be pretty similar to the "true" value (I think it makes more sense to report the true value, with the bootstrapped confidence intervals)
bsOut <- bootEst(bs_time.dogyes, bs_time.dogno)
colMeans(bsOut) 

# now calculate the 95% confidence interval - this takes a little while to run
bs <- as.vector(bsOut[,2])
(bsCI_inout <- bootCI(Dhats[2], bs)) 
```
