---
title: "foxlight_model_selection"
author: "Lara & Kaitlyn"
date: "7/1/2020"
output: html_document
---

# Part 1) Setting up

### 1.1 Load Packages
```{r}
library(MASS)
library(lme4)
library(ggplot2)
library(dplyr)
library(car)
```

### 1.2 Read in record table CSV
```{r record.table}
record.table.all <- read.csv("record_table_0min_deltaT_2020-06-06.csv")
```


### 1.3 create new column for cameras by taking the folder name and reducing to first three characters
```{r new.columns}
record.table.all$Camera <- strtrim(record.table.all$Camera_folder, 5)

record.table.all$Phase <- substr(record.table.all$Camera_folder, 7, 12)

record.table.all$Foxlight <- substr(record.table.all$Camera_folder, 14, 17)
```


### 1.4 Set the time interval (photos that occured within 15 min will count as one detection)
```{r fifteen.min}
record.table.15min <- record.table.all[record.table.all$delta.time.secs >= 900 |  # because 900 seconds = 15 minutes
                                             record.table.all$delta.time.secs == 0, ]
```

### 1.5 Import additional csv


*metadata.raster is normally distributed. The closer to 0, the less the value deviates from the norm. 
```{r import.csv}
# import camera phase operation dates
cam_operations <- read.csv("phase_operation_dates_2020.csv", header=T)

# reading in metadata

# metadata <- read.csv("MicrohabitatData.csv", header=T)
metadata.raster <- read.csv("Raster_metadata_Lara.csv", header=T) %>% 
  select(-c(Latitude, Longitude))

# scale raster values
metadata.raster <- metadata.raster %>%
  mutate_if(is.numeric, scale)
```

### 1.6 Calculate RAI
```{r RAI}
# calculate RAI for each period

# change camera, phase, and species to factor
record.table.15min$Camera <- as.factor(record.table.15min$Camera)
record.table.15min$Phase <- as.factor(record.table.15min$Phase)
record.table.15min$Species <- droplevels(as.factor(record.table.15min$Species))


# calculate number of observations of each species at each camera in each phase
rai <- record.table.15min %>%
  dplyr::group_by(Species, Camera, Phase, .drop = FALSE) %>%
  dplyr::summarise(Detections = n()) 

rai2 <- record.table.15min %>%
  dplyr::group_by(Species, .drop = FALSE) %>%
  dplyr::summarise(Detections = n())
# merge with record table
rai <- merge(rai, cam_operations)


# calculate RAI
rai$RAI <- rai$Detections / rai$Operation

# remove records where camera was operating for <10 days
for (i in 1:nrow(rai)) {
  if(rai$Operation[i] < 10) {
    rai$Detections[i] <- NA
    rai$RAI[i] <- NA
  } 
}

# join raster metadata
rai <- left_join(rai, metadata.raster)

#### now just calculate RAI at each camera, ACROSS ALL PHASES
cam_operations.acrossphases <- read.csv("camera_operation_dates_2020.csv")
# calculate number of observations of each species at each camera
rai.acrossphases <- record.table.15min %>%
  dplyr::group_by(Species, Camera, .drop = FALSE) %>%
  dplyr::summarise(Detections = n()) 
# merge with record table
rai.acrossphases <- merge(rai.acrossphases, cam_operations.acrossphases)
# calculate RAI
rai.acrossphases$RAI <- rai.acrossphases$Detections / rai.acrossphases$Operation
# join with metadata
rai.acrossphases <- left_join(rai.acrossphases, metadata.raster, by = "Camera")
```

# Part 2 - Model selection

Make candidate models and evaluate them!

## Null model/hypothesis - only camera as random effect.
AIC = 424.4
```{r}
fit.null <- glmer.nb(Detections ~ offset(log(Operation)) + (1|Camera), data = subset(rai, Species == "CALA"))
summary(fit.null)
```

## Foxlight Family of Models
### Foxlight only
AIC = 425.6
```{r}
fit1 <- glmer.nb(Detections ~ Foxlight + offset(log(Operation)) + (1|Camera), data = subset(rai, Species == "CALA"))
summary(fit1)
```

### Sheep only
AIC = 424.7
```{r}
fit2 <- glmer.nb(Detections ~ Sheep_Presence + offset(log(Operation)) + (1|Camera), data = subset(rai, Species == "CALA"))
summary(fit2)
```

### Phase only
AIC = 431.1
```{r}
fit3 <- glmer.nb(Detections ~ Phase + offset(log(Operation)) + (1|Camera), data = subset(rai, Species == "CALA"), control=glmerControl(optimizer="bobyqa"))
summary(fit3)
```

### Foxlight + Sheep
AIC = 425.7
VIF = 1.001283
```{r}
fit4 <- glmer.nb(Detections ~ Foxlight + Sheep_Presence + offset(log(Operation)) + (1|Camera), data = subset(rai, Species == "CALA"))
summary(fit4)
vif(fit4)
```

### Foxlight * Sheep
AIC = 427.5
VIF = 2.707
```{r}
fit5 <- glmer.nb(Detections ~ Foxlight * Sheep_Presence + offset(log(Operation)) + (1|Camera), data = subset(rai, Species == "CALA"))
summary(fit5)
vif(fit5)
```
### Foxlight + Phase
AIC = 431.4
VIF = 1.05
```{r}
fit6 <- glmer.nb(Detections ~ Foxlight + Phase + offset(log(Operation)) + (1|Camera), data = subset(rai, Species == "CALA"), control=glmerControl(optimizer="bobyqa"))
summary(fit6)
vif(fit6)
```
### Foxlight * Phase
AIC = 435.3 but error???
```{r}
fit7 <- glmer.nb(Detections ~ Foxlight * Phase + offset(log(Operation)) + (1|Camera), data = subset(rai, Species == "CALA"), control=glmerControl(optimizer="bobyqa"))
summary(fit7)
vif(fit7)
```

## Habitat & Foxlights

### Foxlight * Ruggedness.25
AIC = 424.4
```{r}
fit8 <- glmer.nb(Detections ~ Foxlight * rugged25.clean + offset(log(Operation)) + (1|Camera), data = subset(rai, Species == "CALA"), control=glmerControl(optimizer="bobyqa"))
summary(fit8)
vif(fit8)
```

### Foxlight * Fence Distance
AIC = 431.5
```{r}
fit9 <- glmer.nb(Detections ~ Foxlight * rugged9.clean + offset(log(Operation)) + (1|Camera), data = subset(rai, Species == "CALA"))
summary(fit9)
vif(fit9)
```

### Foxlight * Fence Distance
AIC = 428.4
VIF = 1.29
```{r}
fit10 <- glmer.nb(Detections ~ Foxlight * fence.dist.clean + offset(log(Operation)) + (1|Camera), data = subset(rai, Species == "CALA"), control=glmerControl(optimizer="bobyqa"))
summary(fit10)
vif(fit10)
```
### Foxlight * Rugged.25 + Foxlight * Fence Distance
AIC = 422.7
```{r}
fit11 <- glmer.nb(Detections ~ Foxlight * rugged25.clean + Foxlight * fence.dist.clean + offset(log(Operation)) + (1|Camera), data = subset(rai, Species == "CALA"), control=glmerControl(optimizer="bobyqa"))
summary(fit11)
vif(fit11)
```