---
title: "foxlights_data_analysis"
output: html_document
---
# Part 1) Setting up

### 1.1 Load Packages
```{r library.packages, include = FALSE}
library(nlme)
library(ggplot2)
library(overlap)
library(maptools)
library(lubridate)
library(plyr)
library(camtrapR)
library(dplyr)
library(lme4)
library(MASS)
```

### 1.2 Read in record table CSV
```{r record.table}
record.table.all <- read.csv("record_table_0min_deltaT_2020-06-06.csv")
```


### 1.3 create new column for cameras by taking the folder name and reducing to first three characters
```{r new.columns}
record.table.all$Camera <- strtrim(record.table.all$Camera_folder, 5)

record.table.all$Phase <- substr(record.table.all$Camera_folder, 7, 12)

record.table.all$Foxlight <- substr(record.table.all$Camera_folder, 14, 17)
```


### 1.4 Set the time interval (photos that occured within 15 min will count as one detection)
```{r fifteen.min}
record.table.15min <- record.table.all[record.table.all$delta.time.secs >= 900 |  # because 900 seconds = 15 minutes
                                             record.table.all$delta.time.secs == 0, ]
```

### 1.5 Import additional csv


*metadata.raster is normally distributed. The closer to 0, the less the value deviates from the norm. 
```{r import.csv}
# import camera phase operation dates
cam_operations <- read.csv("phase_operation_dates_2020.csv", header=T)

# reading in metadata

# metadata <- read.csv("MicrohabitatData.csv", header=T)
metadata.raster <- read.csv("Raster_metadata_Lara.csv", header=T)
```

### 1.6 Plotting Species Richness across cameras by lat and long
```{r species.richness.plot}
detectionMaps(CTtable = cam_operations, 
              recordTable = record.table.15min, 
              stationCol = "Camera", 
              Xcol = "Latitude", 
              Ycol = "Longitude",
              richnessPlot = TRUE,
              printLabels = TRUE)

# remove species that we don't care about (here, removing birds, opposum, nothing, mice, pigs, and squirrels) 
for (species in c("2", "Bird", "DIVI", "Ghosts", "Mice", "SCOC", "SCGR", "SCNI")) {
  record.table.15min <- record.table.15min[record.table.15min$Species != species, ]
}
```



### RAI
```{r RAI}
# calculate RAI for each period

# change camera, phase, and species to factor
record.table.15min$Camera <- as.factor(record.table.15min$Camera)
record.table.15min$Phase <- as.factor(record.table.15min$Phase)
record.table.15min$Species <- droplevels(as.factor(record.table.15min$Species))


# calculate number of observations of each species at each camera in each phase
rai <- record.table.15min %>%
  dplyr::group_by(Species, Camera, Phase, .drop = FALSE) %>%
  dplyr::summarise(Detections = n()) 

rai2 <- record.table.15min %>%
  dplyr::group_by(Species, .drop = FALSE) %>%
  dplyr::summarise(Detections = n())
# merge with record table
rai <- merge(rai, cam_operations)


# calculate RAI
rai$RAI <- rai$Detections / rai$Operation

# remove records where camera was operating for <10 days
for (i in 1:nrow(rai)) {
  if(rai$Operation[i] < 10) {
    rai$Detections[i] <- NA
    rai$RAI[i] <- NA
  } 
}

# join raster metadata
rai <- left_join(rai, metadata.raster)

#### now just calculate RAI at each camera, ACROSS ALL PHASES
cam_operations.acrossphases <- read.csv("camera_operation_dates_2020.csv")
# calculate number of observations of each species at each camera
rai.acrossphases <- record.table.15min %>%
  dplyr::group_by(Species, Camera, .drop = FALSE) %>%
  dplyr::summarise(Detections = n()) 
# merge with record table
rai.acrossphases <- merge(rai.acrossphases, cam_operations.acrossphases)
# calculate RAI
rai.acrossphases$RAI <- rai.acrossphases$Detections / rai.acrossphases$Operation
# join with metadata
rai.acrossphases <- left_join(rai.acrossphases, metadata.raster)
```

# Question 1: How do carnivore detections vary by foxlight presence?

## Test if RAI changes between foxlight yes and foxlight no, while controlling for camera

Coyotes
```{r q1.coyotes}
ggplot(subset(rai, Species == "CALA"), aes(x = Foxlight, y = RAI, fill = Foxlight)) +
  geom_boxplot() +
  geom_jitter() +
  labs(title="Coyote Presence With and Without Foxlights",
       x ="Foxlight", y = "Daily Average Number of Coyotes Detected")

# facet_wrap by phase
ggplot(subset(rai, Species == "CALA"), aes(x = Foxlight, y = RAI, fill = Foxlight)) +
  geom_boxplot() +
  facet_wrap(~ Phase)
  
# linear model relating foxlights to rai (instead of a paired t-test; enables us to control for having 4 observations per camera)
fit <- lme(RAI ~ Foxlight, random = ~ 1|Camera, data = subset(rai, Species == "CALA"), na.action = na.omit)
summary(fit)

# try an alternative version with negative binomial - accounts for non-normal data - AIC = 425.6
fit2 <- glmer.nb(Detections ~ Foxlight + offset(log(Operation)) + (1|Camera), data = subset(rai, Species == "CALA"))
summary(fit2)

# null model - AIC = 424.4 (BEST MODEL)
fit3 <- glmer.nb(Detections ~ offset(log(Operation)) + (1|Camera), data = subset(rai, Species == "CALA"))
summary(fit3)

# with camera interaction term, and no random effect - AIC = 428.7
fit4 <- glm.nb(Detections ~ Foxlight*Camera + offset(log(Operation)), data = subset(rai, Species == "CALA"))
summary(fit4)
```

# Question 2: How do coyote detections vary by foxlight presence AND camera trap site? (aka microhabitat)


Coyotes
```{r}
# Boxplot
ggplot(subset(rai, Species == "CALA"), aes(x = Camera, y = RAI, fill = Foxlight)) +
  geom_boxplot()

# Point/Line (each point represents a phase)
ggplot(subset(rai, Species == "CALA"), aes(x = Camera, y = RAI, col = Foxlight)) +
  geom_point() + geom_line() +
  labs(title = "Coyote RAI by Foxlight Presence and Camera Site")

# Line Graph for Coyote RAI by Foxlight Presence and Camera Site
ggplot(subset(rai, Species == "CALA"), aes(x = Phase, y = RAI, col = Foxlight, group = 1)) +
  geom_point() + geom_line(color="black") +
  facet_wrap(~ Camera) +
  scale_x_discrete(label=function(x) abbreviate(x, minlength=1)) +
  labs(title = "Coyote RAI by Foxlight Presence and Camera Site") 
```

Testing Microhabitat Effects
```{r}
# Coyote by grassland + woodland
ggplot(subset(rai, Species == "CALA"), aes(x = Camera, y = RAI, col = Foxlight)) +
  geom_point() + geom_line() +
  facet_wrap(~vegetation.coarser.clean, scales = 'free') +
  scale_x_discrete(label=function(x) abbreviate(x, minlength=3)) +
  labs(title = "Coyote RAI at Camera Trap Sites in Grassland Areas and Camera Trap Sites in Woodland Areas")
```




# Question 3: Carnivore Detections by Sheep Presence 

```{r sheep stuff}
# Camera_operations, add column for sheep presence,

fit <- lme(RAI ~ (Foxlight * Sheep_Presence), random = ~ 1|Camera, data = subset(rai, Species == "CALA"), na.action = na.omit)
summary(fit)

# First way of representing this data 
ggplot(subset(rai, Species == "CALA"), aes(x = Sheep_Presence, y = RAI, fill = Foxlight)) +
  geom_boxplot() +
  geom_jitter() +
  labs(title="Coyote Presence With and Without Sheep Presence and Foxlights",
       x ="Sheep Presence", y = "Daily Average Number of Coyotes Detected")

```


# Question 4: How does foxlight presence impact the daily activity of carnivores? 

Setting up
```{r q4.setup}
# specify date format
record.table.15min$Date <- as.POSIXct(record.table.15min$Date)

# scale clock time to solar time
coords <- matrix(c(-123.079, 39.0013), nrow=1) # note it is c(longitude, latitude)
Coords <- sp::SpatialPoints(coords,
                            proj4string=sp::CRS("+proj=longlat +datum=WGS84"))

# store time as hms
record.table.15min$Time.Corrected <- hms(record.table.15min$Time)

# convert time from HH:MM:SS to decimal hours (HH.HHHH or whatever)
record.table.15min$Time.Decimal <- record.table.15min$Time.Corrected$hour + record.table.15min$Time.Corrected$minute/60 + record.table.15min$Time.Corrected$second/3600

# scale time so that it is between 0 and 1 by dividing by 24
record.table.15min$Time.Scaled <- record.table.15min$Time.Decimal / 24

# convert scaled time (0 to 1) to radians (0 to 2pi)
record.table.15min$Time.Radians <- record.table.15min$Time.Scaled * 2 * pi

# calculate suntime using function from overlap package, and coordinates and dates as formatted above
record.table.15min$Time.Sun <- sunTime(record.table.15min$Time.Radians, record.table.15min$Date, Coords)

# plot coyote activity with and without foxlights
records.coyote.foxyes <- subset(record.table.15min, Species == "CALA" & Foxlight == "FoxY")
records.coyote.foxno <- subset(record.table.15min, Species == "CALA" & Foxlight == "FoxN")

# define overlap function
multilines2 <-function (A, B, xscale = 24, linetype = c(1, 1), linecol = c("#4575b4", 
                                                                           "red"), n.grid = 128, kmax = 3, adjust = 1, 
                        ...) 
{
  bwA <- getBandWidth(A, kmax = kmax)/adjust
  bwB <- getBandWidth(B, kmax = kmax)/adjust
  if (is.na(bwA) || is.na(bwB)) 
    stop("Bandwidth estimation failed.")
  xsc <- if (is.na(xscale))
    1
  else xscale/(2 * pi)
  xxRad <- seq(0, 2 * pi, length = n.grid)
  xx <- xxRad * xsc
  densA <- densityFit(A, xxRad, bwA)/xsc
  densB <- densityFit(B, xxRad, bwB)/xsc
  densOL <- pmin(densA, densB)
  ylim <- c(0, max(densA, densB))
  plot(0, 0, type = "n", ylim = ylim, xlim = range(xx), xlab = "Time", 
       ylab = "Density", xaxt = "n", ...)
  if (is.na(xscale)) {
    axis(1, at = c(0, pi/2, pi, 3 * pi/2, 2 * pi), labels = c("0", 
                                                              expression(pi/2), expression(pi), expression(3 * 
                                                                                                             pi/2), expression(2 * pi)))
  }
  else if (xscale == 24) {
    axis(1, at = c(0, 6, 12, 18, 24), labels = c("Midnight", 
                                                 "Sunrise", "Noon", "Sunset", "Midnight"))
  }
  else {
    axis(1)
  }
  lines(xx, densA, lty = linetype[1], col = linecol[1])
  lines(xx, densB, lty = linetype[2], col = linecol[2])
  return(invisible(list(x = xx, densityA = densA, densityB = densB)))
}
```

Coyotes
```{r q4.dailyactivity.coyotes}
# plot for coyotes
multilines2(records.coyote.foxyes$Time.Sun, records.coyote.foxno$Time.Sun, main = "Coyotes w/ and w/o Foxlight")
legend('top', c("Yes Foxlight", "No Foxlight"), col = c("#4575b4", "red"), lty=c(1,1), bty='n')

#test significance
# this will return the degree of overlap - area under the shared curve (where 0 is completely different and 1 is completely the same)
(Dhats <- overlapEst(records.coyote.foxyes$Time.Sun, records.coyote.foxno$Time.Sun))
# use Dhat4 (they are different formulas - this one is good for sample sizes >50)

# then resample each one 10,000 times (bootstrapping)
bs_time.foxyes <- resample(records.coyote.foxyes$Time.Sun, 10000)
bs_time.foxno <- resample(records.coyote.foxno$Time.Sun, 10000)

# now calculate the resulting bootstrapped overlap value
# it should be pretty similar to the "true" value (I think it makes more sense to report the true value, with the bootstrapped confidence intervals)
bsOut <- bootEst(bs_time.foxyes, bs_time.foxno)
colMeans(bsOut) 

# now calculate the 95% confidence interval - this takes a little while to run
bs <- as.vector(bsOut[,2])
(bsCI_inout <- bootCI(Dhats[2], bs)) 
# use the line with basic confidence intervals (to be honest I'm not sure what these all mean, they are just different methods for calculating)
```


Bobcats
```{r q4.dailyactivity.bobcat}
# plot bobcat activity with and without foxlights
records.bobcat.foxyes <- subset(record.table.15min, Species == "LYRU" & Foxlight == "FoxY")
records.bobcat.foxno <- subset(record.table.15min, Species == "LYRU" & Foxlight == "FoxN")

# plot for bobcats
multilines2(records.bobcat.foxyes$Time.Sun, records.bobcat.foxno$Time.Sun, main = "Bobcats w/ and w/o Foxlight")
legend('top', c("Yes Foxlight", "No Foxlight"), col = c("#4575b4", "red"), lty=c(1,1), bty='n')

#test significance
# this will return the degree of overlap - area under the shared curve (where 0 is completely different and 1 is completely the same)
(Dhats <- overlapEst(records.bobcat.foxyes$Time.Sun, records.bobcat.foxno$Time.Sun))
# use Dhat4 (they are different formulas - this one is good for sample sizes >50)

# then resample each one 10,000 times (bootstrapping)
bs_time.foxyes <- resample(records.bobcat.foxyes$Time.Sun, 10000)
bs_time.foxno <- resample(records.bobcat.foxno$Time.Sun, 10000)

# now calculate the resulting bootstrapped overlap value
# it should be pretty similar to the "true" value (I think it makes more sense to report the true value, with the bootstrapped confidence intervals)
bsOut <- bootEst(bs_time.foxyes, bs_time.foxno)
colMeans(bsOut) 

# now calculate the 95% confidence interval - this takes a little while to run
bs <- as.vector(bsOut[,2])
(bsCI_inout <- bootCI(Dhats[2], bs)) 
```

# Question 5: Habituation
```{r q5.habituation}
##read in the CSV ##

habituation.by.rai.foxY <- read.csv("habituation.by.rai.foxY.2020.csv", header=T)

##testing for coyotes##
ggplot(habituation.by.rai.foxY, aes(x = Phase, y = CALA.RAI)) +
  geom_boxplot() +
  geom_jitter() +
  labs(title = "Coyote RAI by Active Foxlight Phase",
       x = "First or Second Active Foxlight Phase", y = "Coyote RAI")

# linear model relating foxlights to rai (instead of a paired t-test; enables us to control for having 4 observations per camera)
fit <- lme(CALA.RAI ~ Phase, random = ~ 1|Camera, data = habituation.by.rai.foxY, na.action = na.omit)
summary(fit)

```

